---
output: html_document
---
Coursera - Practical Machine Learning Assignment Writeup
========================================================

##Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geek. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

The goal of your project is to predict the manner in which they did the exercise.
The five classes we are having to predict are:  
1. **A** Exact bicep curl
2. **B** Throwing elbows to the front
3. **C** Lifting the dumbbell only halfway
4. **D** Lowering the dumbbell only halfway
5. **E** Throwing the hip forward

## Loading the necessary packages and data

```{r,echo=TRUE,results='hide',message=F,warning=F,comment=NA}
library(caret)
library(randomForest)
train <- read.csv("pml-training.csv", na.strings=c("NA",""), strip.white=T)
test <- read.csv("pml-testing.csv", na.strings=c("NA",""), strip.white=T)
```

## Pre-processing

Lot of columns have empty data like 'NA' or '""'. Let's remove them. Also, remove columns without 
predictive power like user_name, new_window, num_window, raw_timestamp_part_1, raw_timestamp_part_2, 
cvtd_timestampr.
```{r,echo=TRUE,message=F,warning=F,comment=NA}
isNA <- apply(train, 2, function(x) { sum(is.na(x)) })
training <- subset(train[, which(isNA == 0)], 
                    select=-c(X, user_name, new_window, num_window, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp))
isNA <- apply(test, 2, function(x) { sum(is.na(x)) })
testing <- subset(test[, which(isNA == 0)], 
                    select=-c(X, user_name, new_window, num_window, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp))
dim(training)
dim(testing)
```        

## Training a Random Forest Model

Given that the problem is a high-dimensional classification problem
with number of observations much exceeding the number of predictors, random
forest seems like a good choice.

```{r}
set.seed(12345)
model <- randomForest(classe ~ ., data = training)
```

```{r}
model
```

OOB estimate of  error rate: 0.29% looks excellent. The confusion matrix also looks excellent.
Training set has lots of observations so model fits very well. If we had more data to train,  
model training would become time intensive so we could switch to parallel using doMC library 
and/or using less variable according to variable importance.
Let's look at the variable importance.

```{r}
imp <- varImp(model)
imp$Variable <- row.names(imp)
imp[order(imp$Overall, decreasing = T),]
```

```{r eval=F}
predict(model, testing)
```

This simple, almost basic model achieves the perfect 100% accuracy on the testing set.

## Conclusion

Model is perfect for the given testing data so any further analysis is not necessary. 


